Software for the analysis and visualization of deep mutational scanning data

### 2015-03-24 paper/BMC_Bioinformatics_2015/ResponseToReviewers.tex

\begin{response}
This is a \emph{hugely} important point! The first paragraph of the subsection in the \textsl{Background} entitled \textsl{The goal: inferring site-specific amino-acid preferences} now discusses this issue in detail. Essentially, assuming the length $L$ of the gene is $\gg 1$ (which is the case for any reasonable gene), then if the mutant library contains a Poisson distribution of the number of mutations per gene (as is often the case), the problem noted by the reviewer will not arise in practice. The reason is that for such genes, knowing that there is a mutation at site $r$ does not appreciably change the marginal distribution of the number of mutations at other sites (it only changes it by a factor of $\frac{L - 1}{L}$, which in general will be very small). On the other hand, if the mutant library is created such that it contains genes with exactly one mutation per gene (which is now sometimes done), then the problem noted by the reviewer will arise even for realistic genes with $L \gg 1$. The subsection \textsl{The goal: inferring site-specific amino-acid preferences} discusses this issue, and how the problem can be corrected by appropriately defining the frequencies for the wildtype character.
\end{response}

3. Given the substantial increase in runtime and computational complexity of the Bayesian/MCMC approach relative to the simple ratio approach (hours vs fraction of a second), it might be helpful to provide a bit more intuition about what is gained. Is it simply a ``regularization'' where high ratios obtained from low counts are pushed back towards the prior? If so, does the framework really provide more than what could be obtained with a much simpler regularization heuristic?

4. It’s not clear whether \texttt{dms\_inferprefs} outputs the probability of $\phi_{r,x} < 1$ (or $\phi_{r,x} \ne 0$). This is likely to be a useful feature.

5. It's not clear whether the probability values for $\Delta\pi_{r,x} > 0$ or $< 0$ for \texttt{dms\_inferdiffprefs} should be interpreted as nominal or corrected. Providing false discovery rate estimates might be helpful for interpretation (also applies to 4.).

\begin{response}
The posterior probabilities $\Pr\left(\Delta\pi_{r,x} > 0 \mid \mbox{data}\right)$ reported by \texttt{dms\_inferdiffprefs} are \emph{not} corrected. However, note that these are Bayesian posterior probabilities, not classical $P$-values. By default, \texttt{dms\_inferdiffprefs} places a prior over the $\Delta\pi_{r,x}$ values that biases them towards zero (see Equation 31 of the manuscript). Therefore, it is not clear that these posterior probabilities need to be corrected -- there are arguments for why Bayesian posterior probabilities inferred with appropriate priors do not need to be corrected (see \url{http://www.stat.columbia.edu/~gelman/research/published/multiple2f.pdf}). Rather than discuss these complex and incompletely resolved statistical questions, we have simply added the following sentence:
\begin{quote}
In addition, \texttt{dms\_inferdiffprefs} also creates text files that give the posterior probability that $\Delta\pi_{r,x} > 0$ or $< 0$ (these posterior probabilities are \emph{not} corrected to account for the fact that multiple sites are typically being examined, although be default the inferences are made using the regularizing prior in Equation 31).
\end{quote}
\end{response}

6. The function of \textsl{dms\_merge} isn't explained in the paper. Is it simply averaging the two replicate profiles? Could additional statistical power be gained by explicitly modeling replicates within the Bayesian framework?

@@

Quality of written English: Acceptable

Statistical review: Yes, and I have assessed the statistics in my report.

\subsection*{Reviewer \#2}

The author describes a Python package, \texttt{dms\_tools}, which is used to infer character preferences using deep mutational scanning data. This software package allows the user to utilize the method described in Bloom 2014 (\textit{Mol Biol Evol}).

### 2015-03-24 paper/BMC_Bioinformatics_2015/bmc_bioinformatics_article.tex
Responded to reviewer 1 comment 6

\subsubsection*{Implementation.}
The program \texttt{dms\_inferdiffprefs} in the \texttt{dms\_tools} package infers the differential preferences by performing MCMC over the posterior defined by the product of the likelihoods and priors in Equations \ref{eq:pr_nrerr}, \ref{eq:pr_nrstart}, \ref{eq:pr_nrs1}, \ref{eq:pr_nrs2}, \ref{eq:pr_controlpir}, \ref{eq:pr_xir}, \ref{eq:pr_frstart}, and \ref{eq:pr_deltapi}. The MCMC is performed as described for the preferences, and characters can again be any of nucleotides, amino acids, or codons. The program \texttt{dms\_logoplot} visualizes the posterior mean differential preferences via an extension to \texttt{weblogo}~\cite{crooks2004}. In addition, \texttt{dms\_inferdiffprefs} also creates text files that give the posterior probability that $\Delta\pi_{r,x} > 0$ or $< 0$ (in other words, the probability that there is differential selection).

>>>>

\subsubsection*{Implementation.}
The program \texttt{dms\_inferdiffprefs} in the \texttt{dms\_tools} package infers the differential preferences by performing MCMC over the posterior defined by the product of the likelihoods and priors in Equations \ref{eq:pr_nrerr}, \ref{eq:pr_nrstart}, \ref{eq:pr_nrs1}, \ref{eq:pr_nrs2}, \ref{eq:pr_controlpir}, \ref{eq:pr_xir}, \ref{eq:pr_frstart}, and \ref{eq:pr_deltapi}. The MCMC is performed as described for the preferences, and characters can again be any of nucleotides, amino acids, or codons. The program \texttt{dms\_logoplot} visualizes the posterior mean differential preferences via an extension to \texttt{weblogo}~\cite{crooks2004}. In addition, \texttt{dms\_inferdiffprefs} also creates text files that give the posterior probability that $\Delta\pi_{r,x} > 0$ or $< 0$ (these posterior probabilities are \emph{not} corrected to account for the fact that multiple sites are typically being examined, although be default the inferences are made using the regularizing prior in Equation \ref{eq:pr_deltapi}).
